---
title: "Assignment 4 - Part 2"
output: html_notebook
---

**Grade: 51/60**

10/10 Libraries at top \
5/10 Header and description, code chunks named \
10/10 Import code in nb \
0/4 with clean_names() (Extra credit) \
10/10 with assert statement(s) \
10/10 cleaning code (type conversion, NA recoding, other) \
3/6 with description of rationale \
3/4 with tidyverse/dplyr code \

# Part 2


Choose a dataset (this can be the same dataset as the last assignment, or a new one). Define a predictive modeling problem. Create the appropriate files and notebooks to do the following:

1. Describe the data and the problem

2. Read in and check data

2. Clean up the data. 

*Note: You may request that score for these sections replace your score for the coding portion of Assignment 3.*

4. Using best practices, write four functions which add engineered features to the dataset, including a description and rationale. Include a test to determine whether the features are working as intended. 

5. Prepare the data for modeling

Note: this will form the basis for your midterm project. Take advantage of TA's and my office hours. We can provide feedback and guidance. 

---

I have chosen a dataset that contains information on every known satellite orbitting the Earth. 

This is a database compiled by the Union of Concerned Scientists. It contains information on more than 2000 live satellites. The dataset is updated "roughly quarterly." The following information is available in the dataset:

* Name of Satellite
* Country of origin
* Country of operator
* Operating organization
* Users
* Purpose
* Orbit type
* Orbit class
* Information about orbit (eccentricity, inclination etc)
* Date of Launch + Other launch details (location, launch vehicle etc.)
* Expected lifetime
* Contractor (with country)

```{r install_packages}
library(dplyr)
library(janitor)
library(readxl)
library(tidyverse)
library(ggplot2)
library(ggalluvial)
library(tidyr)
library(assertr)
library(lubridate)
```

```{r data_exploration}
satellite_info <- read_excel("UCS_Satellite_Database_4-1-2019.xlsx")
names(satellite_info) <- str_replace_all(names(satellite_info), c(" " = "."))
satellite_info <- satellite_info %>% select(`Name.of.Satellite,.Alternate.Names`:Source...35)
names(satellite_info)
satellite_info %>% clean_names()
#dim(satellite_info)
#typeof(satellite_info)
#str(satellite_info)
```

In the above code chunk, I read in and edit the dataset to convert it into a more usable format. I replaced the spaces in the column names with periods (".") and selected only the columns with information in them, deleting the empty columns in the dataset. The dataset has dimensions of 2062X35, ie, it has 2062 observations over 35 variables. Most of the information is in character format, while the technical information is either numeric or doubles. 

Now, we can write assert functions to check the dataset and clean it.

```{r assert}
is.date <- function(x) {
  !all(is.na(as.Date(as.character(x),format="%y/%m/%d")))
}

satellite_info %>% assert(is.date(satellite_info$Date.of.Launch))
satellite_info %>% assert(is.numeric(satellite_info$`Perigee.(km)`))
satellite_info %>% assert(is.numeric(satellite_info$`Apogee.(km)`))
satellite_info %>% assert(is.numeric(satellite_info$Eccentricity))
```

Once the data has been cleaned and verified, we can start creating engineered features to add to the dataset. 

The first engineered featured I will create will allow for the standardization of the values of information regarding the orbit characteristics such as Eccentricity and Period. 
Below, I create a function called scale2 that standardizes the values given to it around a mean value of 0. I then use a scoped mutate to apply the function to all observations in the Eccentricty and Period.(minutes) columns, and create new columns that contain the standardized values. This will help pick out any outliers and any possible clusters that form when looking at that particular data values. Based on that, some information could be extracted to help evaluate the purpose of the satellite given that information. 

```{r standardization}
scale2 <- function(x, na.rm = FALSE) (x - mean(x, na.rm = na.rm)) / sd(x, na.rm = na.rm)
sat_info <- satellite_info %>% 
  mutate_at(c("Eccentricity", "Period.(minutes)"), list(std = scale2), na.rm = TRUE) %>% 
  select(contains("std"), everything())
```

The second engineered feature that I am adding to this dataset is a column that uses the information given in the Date.of.Launch column and provides the day of the week the launch took place. This is done using the wday function from the lubridate library. The created function dayoflaunch will be applied to the given column using a scoped mutate, to create a new column called Launchday.

```{r date_to_day}
dayoflaunch <- function(date) {
  dayoflaunch <- wday(date)
  dayoflaunch
}

dayoflaunch(today())

sat_info <- sat_info %>% 
  mutate_at(c("Date.of.Launch"), list(Launchday = dayoflaunch)) %>% 
  select(contains("day"), everything())
```

Upon some preliminary examination of the distribution of the Perigee of the orbit of each satellite, it can be seen that the values tend to cluster in three different ranges; less than 10,000, between 10,000 and 30,000 and greater than 30,000. I use this information to divide these values into categories of high, medium and low in the function perigee_type. This is applied to the perigee column (or any other column) using scoped mutates. 

```{r perigee_type}
hist(satellite_info$`Perigee.(km)`)

perigee_type <- function(x, na.rm = TRUE) {
  ifelse(x < 10000, "Low", ifelse(x > 10000 & x < 30000, "Medium", "High"))
}

perigee_type(9999)
perigee_type(10001)
perigee_type(65000)

sat_info <- sat_info %>% 
  mutate_at(c("Perigee.(km)"), list(ptype = perigee_type)) %>% 
  select(contains("ptype"), `Perigee.(km)`, everything())
```

Similarly, inclination can be divded into high or low based on the observed distribution from the histogram. 

```{r inclination_type}

hist(satellite_info$`Inclination.(degrees)`)

inclination_type <- function(x, na.rm = TRUE) {
  ifelse(x < 50, "Low", "High")
}

inclination_type(45)
inclination_type(55)

sat_info <- sat_info %>% 
  mutate_at(c("Inclination.(degrees)"), list(itype = inclination_type)) %>% 
  select(contains("itype"), `Inclination.(degrees)`, everything())

```

Now, with all these features, the dataframe sat_info can be used further for modeling purposes. 